Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-14 shanghai.tex
Historical context and acknowledgements

Quasicharacter sheaves of $G$ form a category, denoted by $\QC(G)$. In
fact, $\QC(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of quasicharacter sheaves on $G$ form a group, denoted by
$\QCiso{G}$; this is the Picard group of the category of
quasicharacter sheaves on $G$. Our application of quasicharacter sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on finding the Picard group of quasicharacter sheaves on $G$, 
for all smooth commutative group schemes $G$, locally of finite type over $\Fq$, 
for which the geometric component group is finitely generated.

We now say a few words about the group schemes 
that appear in our geometrization of admissible
characters of algebraic tori over local fields. Let $K$ be a
non-Archimedean local field with residual field $\Fq$; we place no
restrictions on the characteristic of $K$ -- in particular, it may be
$0$ or not. Write $\OK$ for ring of integers of $K$ and $\pK$ for the
maximal ideal in $\OK$.  Let $T$ be an algebraic torus defined over
$K$. Let $\TT$ be a N\'eron model for $T$; recall that this is a
smooth group scheme locally of finite type over $\OK$. For each
positive integer $d$, let $\GN{T}_d$ be the Greenberg transform of
$\TT \times_{\Spec{\OK}} \Spec{\OK/\pK^d}$; the commutative group scheme $\GN{T}_d$
plays the r\^{o}le of $G$, above.

The main result of this paper is a canonical isomorphism
\[
\QCiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of quasicharacter sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth less than $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasicharacters of
%$K^\times$: the monoidal category of quasicharacter sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

In this paper we show that the monoidal category of quasicharacter sheaves
on the Greenberg transform of the locally finite type Neron model of
$T$ provides an elegant, canonical model for the group of
quasicharacters of $T(K)$, where $K$ is any non-Archimedean local
field and $T$ is any algebraic torus over $K$.

Acknowledgements: Masoud Kamgarpour, Pramod Achar, Hadi Salmasian.

>>>>

We want to situate the choice of term ``quasicharacter sheaf'' in a historical context.
Analogous sheaves on connected, commutative algebraic groups were first studied
by Deligne \cite{SGA4.5}, and more recently Drinfeld has referred to such objects
as character sheaves.  Our quasicharacter sheaves evolved from this notion, and Theorem \ref{thm:snake}
is a direct generalization of the sheaf-function dictionary to the case of commutative
group schemes which are not necessarily connected or finite type.  The other main use
of the term ``character sheaf'' traces back to Lusztig \cite{}.  Most of his energy is devoted to
handling non-commutative groups and though he does define character sheaves for some
non-connected groups, he restricts his attention to reductive groups with cyclic component group.
For extensions of $\ZZ/n\ZZ$ by a torus, it is not difficult to relate his Frobenius-stable character
sheaves to our quasicharacter sheaves.

We close the introduction with a more detailed description of the structure of this paper.

\subsection*{Acknowledgements}

We would like to thank Masoud Kamgarpour, Pramod Achar, and Hadi Salmasian
for allowing us to hijack our Research in Teams meeting at BIRS into a discussion of
quasicharacter sheaves.  Their knowledge and help have been invaluable.

We also gratefully acknowledge the financial support of PIMS and NSERC,
and the hospitality of BIRS during our weeklong stay in May 2012.
\David{Should these be abbreviated or written out in full?}
